# Clean

## About

This page documents the process of cleaning the 2010-2025 ARRL Field Day data.

The datasets have different field names across the years, so the field names must be normalized in order to combine the datasets for visualization.

## Tidy names

```{r}
clean_fd_names_save <- function(fn, clean_file_dir = "data/cleaned") {
  assertthat::is.string(fn)
  assertthat::is.readable(fn)
  
  df <- readr::read_csv(fn, show_col_types = FALSE)
  assertthat::assert_that(tibble::is_tibble(df))
  
  # Clean names
  df_clean <- janitor::clean_names(df)

  fn_no_ext <- basename(tools::file_path_sans_ext(fn))
  fn <- file.path(clean_file_dir, paste0(fn_no_ext, "-clean-names.csv"))
  readr::write_csv(df_clean, fn)
  fn
}
```

Clean all names for all files in `data/raw`.

```{r}
fl <- list.files("data/raw", full.names = TRUE)

purrr::map(fl, clean_fd_names_save)
```

## Exploration

Let's explore the column names to see which ones represent the same data.

```{r}
show_names <- function(fn) {
  assertthat::is.string(fn)
  assertthat::is.readable(fn)
  
  df <- readr::read_csv(fn, show_col_types = FALSE)
  assertthat::assert_that(tibble::is_tibble(df))
  file_year <- stringr::str_extract(fn, pattern = "[0-9]{4}")
  sprintf("%s\n", paste0("--- ", file_year, " ---"))
  names(df)
}

fl <- list.files("data/cleaned", pattern = "clean-names", full.names = TRUE)

purrr::map(fl, show_names)
```

```{r}
fd_dfs <- purrr::map(fl, readr::read_csv)
janitor::compare_df_cols(fd_dfs)
```

## Rename by variable and year

```{r}
fd2010 <- readr::read_csv(file = "data/cleaned/field-day-2010-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2010)

fd2010 <- fd2010 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2010, file = "data/cleaned/field-day-2010-normed.csv")
```

```{r}
fd2011 <- readr::read_csv(file = "data/cleaned/field-day-2011-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2011)

fd2011 <- fd2011 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2011, file = "data/cleaned/field-day-2011-normed.csv")
```

```{r}
fd2012 <- readr::read_csv(file = "data/cleaned/field-day-2012-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2012)

fd2012 <- fd2012 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2012, file = "data/cleaned/field-day-2012-normed.csv")
```

```{r}
fd2013 <- readr::read_csv(file = "data/cleaned/field-day-2013-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2013)

fd2013 <- fd2013 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2013, file = "data/cleaned/field-day-2013-normed.csv")
```

```{r}
fd2014 <- readr::read_csv(file = "data/cleaned/field-day-2014-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2014)

fd2014 <- fd2014 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2014, file = "data/cleaned/field-day-2014-normed.csv")
```

```{r}
fd2015 <- readr::read_csv(file = "data/cleaned/field-day-2015-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2015)

fd2015 <- fd2015 |>
  dplyr::rename("section" = "sec") |>
  dplyr::mutate(transmitters = stringr::str_extract(class, pattern = "[0-9]+"))

readr::write_csv(fd2015, file = "data/cleaned/field-day-2015-normed.csv")
```

```{r}
fd2016 <- readr::read_csv(file = "data/cleaned/field-day-2016-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2016)

fd2016 <- fd2016 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                class = category,
                transmitters = number_of_transmitters)

readr::write_csv(fd2016, file = "data/cleaned/field-day-2016-normed.csv")
```

```{r}
fd2017 <- readr::read_csv(file = "data/cleaned/field-day-2017-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2017)

fd2017 <- fd2017 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                call = call_does_not_include_checklogs,
                class = category,
                transmitters = number_of_transmitters)

readr::write_csv(fd2017, file = "data/cleaned/field-day-2017-normed.csv")
```

```{r}
fd2018 <- readr::read_csv(file = "data/cleaned/field-day-2018-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2018)

fd2018 <- fd2018 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                class = qst_class_plus,
                transmitters = number_of_transmitters)

readr::write_csv(fd2018, file = "data/cleaned/field-day-2018-normed.csv")
```

```{r}
fd2019 <- readr::read_csv(file = "data/cleaned/field-day-2019-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2019)

fd2019 <- fd2019 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = qst_class,
                number_of_participants = participants,
                transmitters = xmtrs)

readr::write_csv(fd2019, file = "data/cleaned/field-day-2019-normed.csv")
```

```{r}
fd2020 <- readr::read_csv(file = "data/cleaned/field-day-2020-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2020)

fd2020 <- fd2020 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants,
                final_score = score,
                transmitters = xmtrs,
                power_multiplier = pwr_mult)

readr::write_csv(fd2020, file = "data/cleaned/field-day-2020-normed.csv")
```

```{r}
fd2021 <- readr::read_csv(file = "data/cleaned/field-day-2021-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2021)

fd2021 <- fd2021 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants,
                final_score = score,
                transmitters = xmtrs,
                power_multiplier = pwr_mult)

readr::write_csv(fd2021, file = "data/cleaned/field-day-2021-normed.csv")
```

```{r}
fd2022 <- readr::read_csv(file = "data/cleaned/field-day-2022-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2022)

fd2022 <- fd2022 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants)

readr::write_csv(fd2022, file = "data/cleaned/field-day-2022-normed.csv")
```

```{r}
fd2023 <- readr::read_csv(file = "data/cleaned/field-day-2023-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2023)

fd2023 <- fd2023 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants)

readr::write_csv(fd2023, file = "data/cleaned/field-day-2023-normed.csv")
```

```{r}
fd2024 <- readr::read_csv(file = "data/cleaned/field-day-2024-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2024)

fd2024 <- fd2024 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os,
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants)

readr::write_csv(fd2024, file = "data/cleaned/field-day-2024-normed.csv")
```

```{r}
fd2025 <- readr::read_csv(file = "data/cleaned/field-day-2025-clean-names.csv", 
                          show_col_types = FALSE)
names(fd2025)

fd2025 <- fd2025 |>
  dplyr::rename(total_cw_qsos = total_cw_qs_os,
                total_phone_qsos = total_phone_qs_os,
                total_digital_qsos = total_digital_qs_os, 
                total_qsos = total_qs_os,
                class = category,
                number_of_participants = participants)

readr::write_csv(fd2025, file = "data/cleaned/field-day-2025-normed.csv")
```

Let's now check again.

```{r}
fd_dfs <- list(
  fd2010,
  fd2011,
  fd2012,
  fd2013,
  fd2014,
  fd2015,
  fd2016,
  fd2017,
  fd2018,
  fd2019,
  fd2020,
  fd2021,
  fd2022,
  fd2023,
  fd2024,
  fd2025
)

janitor::compare_df_cols(fd_dfs)
```

## Merge cleaned

```{r}
import_add_year <- function(fn) {
  assertthat::is.string(fn)
  assertthat::is.readable(fn)
  
  df <- readr::read_csv(fn, show_col_types = FALSE)
  assertthat::assert_that(tibble::is_tibble(df))
  df$year <- stringr::str_extract(fn, "[0-9]{4}")
  df
}
fl <- list.files("data/cleaned", pattern = "normed", full.names = TRUE)

fd_all_df <- purrr::map(fl, import_add_year) |>
  purrr::list_rbind()

readr::write_csv(fd_all_df, file = "data/aggregate/field-day-2010-2025.csv")
```

