# Gather

## About

This page documents the process of gathering Field Day data from the ARRL website.

## Download

The ARRL has comma-separated value (text) files for Field Day results here: <https://field-day.arrl.org/fdresults.php>.

CSV data are available for 2010-2025.

::: {.callout-note}
## Future work

QST articles go back to 2000, so the results from 2000-2009 could eventually be extracted from these PDF files and turned into actionable data.

:::

Create functions to download the data given year as an input.

```{r}
#| label: specify-helper-function
download_fd_data <- function(year = 2025) {
  assertthat::is.number(year)
  assertthat::assert_that(year > 0)
  
  url <- paste0("https://contests.arrl.org/ContestResults/", year, "/field-day-", year, ".csv")
  readr::read_csv(url, show_col_types = FALSE)
}
```

```{r}
#| label: save-fd-data
save_fd_data <- function(df, year, data_dir = "data/raw", force_create_dir = TRUE) {
  assertthat::assert_that(tibble::is_tibble(df))
  
  assertthat::is.number(year)
  assertthat::assert_that(year > 0)
  
  assertthat::is.string(data_dir)
  assertthat::assert_that(is.logical(force_create_dir))
  if (!assertthat::is.writeable(data_dir)) {
    if (force_create_dir) {
      dir.create(data_dir)
    } else {
      warning(paste0(data_dir, " does not exist. Exiting."))
      return()
    }
  }
  
  fn <- file.path(data_dir, paste0("field-day-", year, ".csv"))
  readr::write_csv(df, fn)
} 
```

```{r}
download_save_df_data <- function(year = 2025) {
  df <- download_fd_data(year)
  save_fd_data(df, year)
}
```

With these functions, we can use the {purrr} package to download and save all of the available data files.

```{r}
#| label: download-save-2010-2025
years <- c(2010:2025)
purrr::map(years, download_save_df_data, .progress = TRUE)
```

