# Visualize

## About

This page documents visualizations of the ARRL Field Day 2010-2025 data.

## Setup

Here I `source` the {ggplot2} library to make it easier to make plots later. I also define some default colors for those plots. Defining these here makes it easier to modify them site-wide later on.

```{r}
library(ggplot2)

max_color = "red"
median_color = "blue"
tot_color = "green"
```

## Import

The gathering and cleaning pages handle those functions. They save a cleaned data file that we reimport here.

::: {.callout-important}
## DRY WIT

A popular adage in computer science and data science is DRY WIT: Don't Repeat Yourself; Write It Down.

I adopt that mantra here by automating each step and by saving the intermediate products.

:::

```{r}
fd <- readr::read_csv("data/aggregate/field-day-2010-2025.csv", show_col_types = FALSE)
```

## Generate summaries

To begin, we create a smaller data summary.

```{r}
fd_by_year <- fd |>
  dplyr::group_by(year) |>
  dplyr::summarize(
    median_score = median(final_score, na.rm = TRUE),
    max_score = max(final_score, na.rm = TRUE),
    tot_score = sum(final_score, na.rm = TRUE),
    median_xmtrs = median(transmitters),
    max_xmtrs = max(transmitters),
    median_particip = median(number_of_participants, na.rm = TRUE),
    max_particip = max(number_of_participants, na.rm = TRUE),
    tot_particip = sum(number_of_participants, na.rm = TRUE),
    median_qsos = median(total_qsos, na.rm = TRUE),
    max_qsos = max(total_qsos, na.rm = TRUE),
    tot_qsos = sum(total_qsos, na.rm = TRUE),
    n_stations = length(unique(call))
  )
```

## Scores by year

This section visualizes Field Day scores across the 2010-2025 period.

::: {.panel-tabset}

### Median

```{r}
#| label: fig-fd-median-score-2010-2025
#| fig-cap: "Median Field Day score 2010-2025"
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = median_score)) +
  geom_line(aes(x = year, y = median_score), color = median_color) +
  ylab("Median score")
```

### Maximum

```{r}
#| label: fig-fd-max-score-2010-2025
#| fig-cap: "Maximum Field Day score 2010-2025"
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = max_score)) +
  geom_line(aes(x = year, y = max_score), color = max_color) +
  ylab("Max score")
```

### Maximum & median

```{r}
#| label: fig-fd-max-median-scores-2010-2015
#| fig-cap: "Maximum (red) and median (blue) Field Day scores 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = max_score), color = max_color) +
  geom_line(aes(x = year, y = median_score), color = median_color) +
  ylab("score")
```

### Total

```{r}
#| label: fig-fd-total-score-2010-2025
#| fig-cap: "Total Field Day scores 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = tot_score), color = tot_color) +
  ylab("Score total")
```

:::

## Participants by year

The following plots show data about the number of participants per station per year.

::: {.panel-tabset}

### Median

```{r}
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = median_particip)) +
  geom_line(aes(x = year, y = median_particip), color = median_color) +
  ylab("Median participants/station")
```

### Maximum

```{r}
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = max_particip)) +
  geom_line(aes(x = year, y = max_particip), color = max_color) +
  ylab("Maximum participants/station")
```

### Max & median

```{r}
#| label: fig-fd-max-median-particip-2010-2015
#| fig-cap: "Maximum (red) and median (blue) Field Day scores 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = max_particip), color = max_color) +
  geom_line(aes(x = year, y = median_particip), color = median_color) +
  ylab("Participants/station")
```

### Total

```{r}
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = tot_particip), color = tot_color) +
  ylab("Total participants")
```

:::

## QSOs by year

::: {.panel-tabset}

### Median

```{r}
#| label: fig-fd-median-qsos-2010-2025
#| fig-cap: "Median Field Day QSOs/station 2010-2025"
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = median_qsos)) +
  geom_line(aes(x = year, y = median_qsos), color = median_color) +
  ylab("Median QSOs/station")
```

### Maximum

```{r}
#| label: fig-fd-max-qsos-2010-2025
#| fig-cap: "Maximum Field Day QSOs/station 2010-2025"
fd_by_year |>
  ggplot() +
  geom_point(aes(x = year, y = max_qsos)) +
  geom_line(aes(x = year, y = max_qsos), color = max_color) +
  ylab("Max QSOs")
```

### Maximum & median

```{r}
#| label: fig-fd-max-median-qsos-2010-2015
#| fig-cap: "Maximum (red) and median (blue) Field Day QSOs/station 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = max_qsos), color = max_color) +
  geom_line(aes(x = year, y = median_qsos), color = median_color) +
  ylab("Max & median QSOs")
```

### Total

```{r}
#| label: fig-fd-total-qsos-2010-2025
#| fig-cap: "Total Field Day QSOs 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = tot_qsos), color = tot_color) +
  ylab("Total QSOs")
```

:::

## Participating stations

This section visualizes data about the participating stations.

```{r}
#| label: fig-fd-n-stations-2010-2025
#| fig-cap: "Number of participating stations 2010-2025"

fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = n_stations), color = tot_color) +
  ylab("Total participating stations")  
```

## Transmitters

```{r}
#| label: fig-fd-max-median-xmtrs-2010-2015
#| fig-cap: "Maximum (red) and median (blue) Field Day transmitters/station 2010-2025"
fd_by_year |>
  ggplot() +
  geom_line(aes(x = year, y = max_xmtrs), color = max_color) +
  geom_line(aes(x = year, y = median_xmtrs), color = median_color) +
  ylab("Max & median transmitters")
```
